---
title: "STAT 325 Final Exam"
author: "Claude Lee"
date: "May 10, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(Lahman)
library(lubridate)
```

# 1. 
### Primary keys for `Master`, `Batting`, `Pitching`, and `Fielding`.
```{r 1-1, collapse = TRUE}
nrow(Master)
count(count(Master, playerID))[[1]]
```

Primary key for `Master`: `playerID`


```{r 1-2, collapse = TRUE}
nrow(Batting)
group_by(Batting, playerID, yearID) %>% 
  nrow()
```

Primary key for `Batting`: `playerID` and `yearID` together.


```{r 1-3, collapse = TRUE}
nrow(Pitching)
group_by(Pitching, playerID, yearID) %>% 
  nrow()
```

Primary key for `Pitching`: `playerID`, and `yearID` together.


```{r 1-4, collapse = TRUE}
nrow(Fielding)
group_by(Fielding, playerID, yearID) %>% 
  nrow()
```

Primary key for `Fielding`: `playerID`, and `yearID` together.




# 2. 
### Cacreer statistics for each player with `Batting`.
```{r 2-1}
Batting2 <- Batting %>%
  group_by(playerID) %>% 
  mutate(G = sum(G, na.rm = TRUE), AB = sum(AB, na.rm = TRUE), 
         R = sum(R, na.rm = TRUE), H = sum(H, na.rm = TRUE), 
         X2B = sum(X2B, na.rm = TRUE), X3B = sum(X3B, na.rm = TRUE),
         HR = sum(HR, na.rm = TRUE), RBI = sum(RBI, na.rm = TRUE), 
         SB = sum(SB, na.rm = TRUE), CS = sum(CS, na.rm = TRUE), 
         BB = sum(BB, na.rm = TRUE), SO = sum(SO, na.rm = TRUE), 
         IBB = sum(IBB, na.rm = TRUE), HBP = sum(HBP, na.rm = TRUE), 
         SH = sum(SH, na.rm = TRUE), SF = sum(SF, na.rm = TRUE),
         GIDP = sum(GIDP, na.rm = TRUE))
as_tibble(Batting2)
```

Use `mutate`.

It was possible to make a function with loops to get sums for multiple variables, however since the time that it takes to run all the loops, it becomes inefficient (takes about 3 minutes). Below is the function that runs the loop.
```{r 2-2, eval=FALSE, include=TRUE}
mutatesum1 <- function(df,keyIndex, i, j) {
  #Arrange the dataframe with key first.
  #keyIndex: position of the key in column
  #i: position of the variable to start
  #j: position of the variable to finish
  varsum <- matrix(nrow = nrow(df), ncol = j-i+1)
  for (obs in 1:nrow(df)) {
    for (vars in i:j) {
      if (is.na(df[obs, vars])) {
        varsum[obs, vars-i+1] <- 0
      } 
      if (obs == 1) {
        varsum[obs, vars-i+1] <- df[obs, vars]
      } else if (df[obs, keyIndex] == df[obs-1, keyIndex]) {
        varsum[obs, vars-i+1] <- df[obs, vars] + varsum[obs-1, vars-i+1]
        l <- 1
        while (df[obs, keyIndex] == df[obs-l, keyIndex]) {
          varsum[obs-l, vars-i+1] <- varsum[obs, vars-i+1]
          l <- l+1
          if (obs-l == 0){
            break
          }
        }
      } else {
        varsum[obs, vars-i+1] <- df[obs, vars]
      }
    }
  }
  df[, i:j] <- varsum
}
```




# 3. 
### Do (2) with `Pitching`.
```{r 3}
Pitching2 <- Pitching %>%
  group_by(playerID) %>% 
  mutate(W = sum(W, na.rm = TRUE), L = sum(L, na.rm = TRUE), 
         G = sum(G, na.rm = TRUE), GS = sum(GS, na.rm = TRUE), 
         CG = sum(CG, na.rm = TRUE), SHO = sum(SHO, na.rm = TRUE),
         SV = sum(SV, na.rm = TRUE), IPouts = sum(IPouts, na.rm = TRUE), 
         H = sum(H, na.rm = TRUE), ER = sum(ER, na.rm = TRUE), 
         HR = sum(HR, na.rm = TRUE), BB = sum(BB, na.rm = TRUE), 
         SO = sum(SO, na.rm = TRUE), IBB = sum(IBB, na.rm = TRUE),
         WP = sum(WP, na.rm = TRUE), HBP = sum(HBP, na.rm = TRUE),
         BK = sum(BK, na.rm = TRUE), BFP = sum(BFP, na.rm = TRUE),
         GF = sum(GF, na.rm = TRUE), R = sum(R, na.rm = TRUE),
         SH = sum(SH, na.rm = TRUE), SF = sum(SF, na.rm = TRUE),
         GIDP = sum(GIDP, na.rm = TRUE))
as_tibble(Pitching2)
```




# 4. 
### `Batting4` with statistics and players' info.
```{r 4}
Batting3 <- as_tibble(battingStats(Batting2))

#selects variables we need from `Master` data.
BattingInfo <- Master %>%
  select(playerID, starts_with("name"), bats, throws)

#join the data on the right side
Batting4 <- Batting3 %>% 
  left_join(Master, by = "playerID")

as_tibble(select(Batting4, playerID, starts_with("name"), bats, throws, everything()))
```




# 5.
### Combine `Batting4` and `Pitching2`.
```{r 5-1}
BatPitch1 <- Batting4 %>% 
  filter(PA > 0) %>% 
  semi_join(Pitching2, by = "playerID") %>% 
  left_join(Pitching2, by = "playerID")
BatPitch1
```

This table has some problem. Some varibles are shared in both dataframe, so it creates duplicated varaibles. So, let's filter out those variables, then join them.

```{r 5.2}
# Defind a function that finds a vector of varaible names that are in both data frame.
dupVar <- function(x,y, from) {
  dupVar <- vector()
  count <- 1
  for (i in from:ncol(x)) {
    if (names(x[i]) %in% variable.names(y)){
      dupVar[count] <- names(x[i])
      count <- 1+count
    }
  }
  return(dupVar)
}

# Remove the variable (By start from 2, the first variable `playerID` is kept).
PitchRM1 <- Pitching2 %>% 
  select(-c(dupVar(Pitching2, Batting4,2)))

# Join
BatPitch1 <- Batting4 %>% 
  filter(PA > 0) %>% 
  semi_join(PitchRM1, by = "playerID") %>% 
  left_join(PitchRM1, by = "playerID")

BatPitch1
```




# 6. 
### Write a function `HR_rate` that calculates `HR/G` and returns the original table with sorted by new variable.
```{r 6}
HR_rate <- function(df) {
  df2 <- df %>% 
    mutate(HR_G = HR / G) %>% 
    select(playerID:lgID, HR_G, everything()) %>% 
    arrange(desc(HR_G))
  return(df2)
}

as_tibble(HR_rate(Batting))
```

In order to keep the `Batting` as it as, we will return a new table with the new variable, instead of overriding the original table, since it gets used later too.


# 7. 
### Create `Batting5` by relevel `lgID` of `Batting` as `AL`, `NL`, and `Other`.

First, let's check the levels of `lgID`.

```{r 7-1, collapse = TRUE}
levels(Batting$lgID)
```

There are `AA`, `FL`, `NA`, `PL`, and `UA` that are not `AL` or `NL`. Therefore, let's collapse them into the `Other` level.

```{r 7-2}
Batting5 <- Batting %>%
  mutate(lgID = fct_collapse(lgID,
                             AL = "AL",
                             NL = "NL",
                             Other = c("AA", "FL", "NA", "PL", "UA")))
as_tibble(Batting5)
```




# 8. 
### Create integer variables `debutYear`, `debutMonth`, and `debutDay` from `Master$debut`. Then, create `debutDate` and `debutAge` into `Master2`.
```{r 8}
Master2 <- Master %>% 
  mutate(debutYear = as.integer(str_sub(debut, 1, 4)), 
         debutMonth = as.integer(str_sub(debut, 6, 7)),
         debutDay = as.integer(str_sub(debut, 9, 10)),
         debutDate = make_date(debutYear, debutMonth, debutDay), 
         birthDate = ymd(birthDate),
         debutAge = as.period(interval(birthDate, debutDate), unit = "year")$year)

as_tibble(select(Master2, playerID, debutDate, debutAge, everything()))
```




# 9. 
### Use `Master$birthCountry` to create a factor variable called `birthContinent` that gives the continent of each player's birth.

Before creating a factor, let's check the levels of the variable.

```{r 9-1, collapse = TRUE}
unique(Master2$birthCountry)
```

Now, let's convert the variable as a factor and define the levels.

```{r 9-2}
Master2 <- Master2 %>% 
  mutate(birthCountry = factor(birthCountry, levels = unique(birthCountry))) %>% 
  mutate(birthContinent = fct_collapse(birthCountry, 
                                       `North America` = c("USA", "CAN"),
                                       `Latin America` = c("D.R.", "Venezuela",
                                                           "Cuba", "Mexico",
                                                           "Colombia", "Nicaragua", 
                                                           "Bahamas", "P.R.", "Curacao", 
                                                           "Aruba", "Jamaica", "Brazil", 
                                                           "Belize", "Honduras", 
                                                           "American Samoa", "V.I."),
                                       `Europe` = c("Germany", "Norway", "Czech Republic",
                                                    "Netherlands", "Ireland", "Italy",
                                                    "United Kingdom", "France", "Sweden", 
                                                    "Spain", "Greece", "Poland", "Denmark", 
                                                    "Switzerland", "Austria", "Belgium", 
                                                    "Finland", "Lithuania", "Latvia", 
                                                    "Slovakia", "Portugal"),
                                       `Asia` = c("Russia", "Japan", "South Korea", 
                                                  "Hong Kong", "Afghanistan", "Taiwan",
                                                  "Philippines", "Viet Nam", "China", 
                                                  "Indonesia", "Saudi Arabia"),
                                       `Oceania` = c("Australia", "Guam", "Singapore"),
                                       `Afrika` = c("South Africa"),
                                       `Other` = c("At Sea")))
as_tibble(select(Master2, playerID, birthCountry, birthContinent, debutAge, debutDate, everything()))
```




# 10.
### `Master2` with checking difference in `bats` and `throws`.
```{r 10}
SameDiff <- function(df) {
  Same_Dif <- character(nrow(df))
  for (i in 1:nrow(df)) {
    if (is.na(as.character(df$bats[i]) == as.character(df$throws[i]))){
      Same_Dif[i] <- NA
    } else if (as.character(df$bats[i]) == as.character(df$throws[i])) {
      Same_Dif[i] <-  "Same"
    } else {
      Same_Dif[i] <-  "Different"
    }
  }
  df2 <- df %>% 
    mutate(Same_Dif = Same_Dif)
  return(df2)
}

as_tibble(select(SameDiff(Master2), playerID, Same_Dif, bats, throws, birthCountry, birthContinent, debutAge, debutDate, everything()))
```




# 11.
### Same as (10), but only with one of the `map*` functions.
```{r 11}
SameDiffFnc <- function(x, y) {
  if (is.na(x==y)) {
    return(NA)
  } else if (x == y) {
    return("Same")
  } else {
    return("Different")
  }
}

SameDiffDf <- function(df) {
  Same_Dif_map <- map2_chr(as.character(df$bats), as.character(df$throws), SameDiffFnc)
  df2 <- df %>% 
    mutate(Same_Dif_map = Same_Dif_map)
  return(df2)
}

as_tibble(select(SameDiffDf(Master2), playerID, Same_Dif_map, bats, throws, birthCountry, birthContinent, debutAge, debutDate, everything()))
```

